---
title: "AIS Tech Lesson - R Intro 1"
author: "AIS UIUC"
date: "August 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**First thing first, set working directory**
- files from working directory can be accessed through file names
- files outside working directory have to be accessed through it's entire directory
  + "C:/Non-Default aka Real Shit/College is a hot mess/Fall 2018/STAT 434/Homework/1"

## R as a calculator

#### Addition
```{r}
# What is 5 plus 7 (what is "#" ?)
```

#### Subtraction
```{r}
# what is 5 minus 7?
```

#### Multiplication
```{r}
# what is 3 times 5? 
```

#### Division
```{r}
# what is 3 divided by 5
```

#### **Exponentiation** 
- the operation of raising one quantity to the power of another - $$x^n$$
```{r}
# what is the square of 9
# what is the cube of 9

# How do I do sqaure root then?
```

#### Order of calculation - same rule as in math, use "()" if confused
```{r}
# What is the order of calculation of this? 
1 + 2^(4/2)*5 
```

#### **Modulo**
```{r}
# Take a guess, what do the following commands return? 
5 %% 2
7 %% 3
22 %% 4
```
Why modulo? ...

### **Built-in Functions**
- Function you can directly call and use without loading a package
- [Here is a link for R's built-in math functions](https://www.rdocumentation.org/packages/raster/versions/2.6-7/topics/Math-methods)

#### Exponentials
```{r}
# Is this number familar to you? 
exp(1)
```

#### logarithmic
```{r}
# base-10 logarithmic
# base-2 logarithmic
# what about natural logarithmic - base is exp(1), aka 2.7818......
# base 5? 
```






## Variable
- **Variable has a name, variable has a value, variable's name represents it's value (all you need to know to continue)**
- Variable is a reference to a point/location of memory in which a specific value is stored
- When referenced, the value stored in memory location is accessed
- Why variable?
  + assign meaning of a value (easier to read code)
  + reuse 
  
  
#### Assigning a value to a variable
```{r}
apple <- 5
peach <- 7

# how many fruit do we have? assign the answer to a new variable 
```
- you can assign to a variable value of many **data types**
- **scope** - will cover when we touch on **functions**





## Data Types
#### Common Data Types
-Numeric
-String/Character
-Categorical/Factorial
  + a variable with a fixed set of values called levels
-Boolean/logical
  + True/False
-Others
```{r}
a <- 12
b <- 12.12
c <- 'twelve' # just a set of "" will do the trick
d <- "12"
e <- TRUE
f <- FALSE
g <- a==a
h <- a==b
i <- factor(c)
j <- NULL # why "NULL"? 
k <- NA
class(k) # "class()" check data type of an varaiable for you
```





#### Conversion
```{r}
as.numeric("12")
as.character(k)
as.factor(f)
```
- **Take-away: Conversion cannot take place in any direction**
- [NULL vs NA](https://www.r-bloggers.com/r-na-vs-null/):
  + "NULL" - often returned by expressions and functions whose values are undefined.
  + "NA" - missing value, can be coerced to any data type, by default a logical

## Logical/Boolean Operator
- return **TRUE/FALSE**
#### Test out the operators
```{r}
1 < 2
1 > 2
2 > 1
1 >= 1 # larger or equal to
1 >= 2
2 >= 1
2 == 2
2 == 1
# "<=" ...... you know what it is 
2 != 2
2 != 1
!TRUE
!(1==1)
!(1!=1) # so what does "!x" do? ......
```
#### multiple conditions
```{r}
x = TRUE
y = FALSE
(x|y) # x or y
(x&y) # 
```
- so what does "or", aka "|", do? ...
- what does "and", aka "&",  do? ...




## Vector Basics
- Definition: A vector is a sequence of data elements of the **same basic type**
- Most important feature - **vectorized operation**

#### Creating a vector
```{r}
# create a vector & check it's type 
a <- c(1,2,3)
class(a)

# how to check if a variable is vector?
is.vector(a)
is.vector(123)
is.vector("one hundred twenty three")
```
- What..the..h.. - Well, vector is R's base data structure, as long as you don't create an variable explicitly as other structure, it is a vector

#### Coercion
```{r}
a <- c(1, "one", TRUE) # multiple data types, that's impossible by definition! 
a
class(a)
```
- Coecion ordering, lower canbe coerced into higher (of only the data types we touched here)
  + logical < numeric < factor < character
  + Question, what is the numeric conversion of **TRUE/FALSE**? 

#### Vectorized Operation
- Vectorized operation is what makes vector a great data structure, let's see this in action
```{r}
# create a vector, name it whatever, assign it with 6 numeric values(your pick)

# add a numeric value to it (any numeric value)

# crate 3 other vector, one with 3 numeric values, one with 6 numeric values, and one with 5 numeric values

# add each of these to your to your first whatever vector

```

-Try it on another data type
```{r}
# What does the following code do?
logics <- rep(TRUE, 15)
# print(logics)
logics <- !logics
# print(logics)
booleans <- c(FALSE, TRUE, FALSE)
booleans & logics
booleans | logics
```

- Not just on operations through operators, through functions, too
```{r}
this <- sample.int(100, 10)
print(this)
that <- exp(this)
print(that)
```


#### Indexing/slicing vectors
```{r}
a <- seq(0, 10, by=0.1)
a

# how to access the 50-th element?

# how to access the first 15 element?

# how to access the 16th to 30th element?

a[1:10]
```


#### Aggregation functions over numeric vectors
- Aggregation of vector - to collapse a multi-element vector into a single value (average, sum, etc.)
```{r}
a <- c(123, 1232, 123124, 123421, 12, 1.12)
# what is the mean(average) of a
# median
# standard deviation
# sum
# minimum
# maximum
summary(a)
```

#### [*Other data structure*](http://adv-r.had.co.nz/Data-structures.html)








## Base Plot

#### Histogram
```{r}
# generate a set of random numbers from standard normal distribution
ran <- rnorm(1000)
hist(ran, breaks = 30, col = 'light blue')
curve(dnorm(x, mean=0, sd=1), add = TRUE, col = 'red')
```
- Useful base plot arguments (for this stage)
  + main - title of the plot
  + xlab - label for x axis
  + ylab - label for y axis
  + xlim - limit for x axis
  + ylim - limit for y axis

#### barplot
```{r}
# creating the data, don't worry about the details here
count <- c(15, 36, 40, 78)
fruits <- c("apple", "pear", "peach", "banana")

barplot(count, names.arg = fruits)
```


#### Scatter plot
```{r}
# using r built-in data mtcars
attach(mtcars)
plot(x = wt, y = mpg, col=cyl)

# now customize this plot to make sense
```


## Getting Help 
#### in R/RStudio
```{r}

```
#### Google, Stackoverflow and more

## Reading a csv file
```{r}

```


## Data Frame

